Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision

### 2023-04-17 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
prettify the slides again

---
title: "Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision"
subtitle: ""  
author: 
  - "Weihao (Patrick) Li"
date: '`r format(Sys.Date(), format="%B %d, %Y")`'
output:
  xaringan::moon_reader:
    seal: false
    css: [xaringan-themer.css, mine.css]
    nature:
      countdown: 55000
      beforeInit: "scripts/my_macro.js"
      countIncrementalSlides: false
      slideNumberFormat: "%current%/%total%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
---

@@

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(tidyverse)
library(visage)
style_solarized_light(
  background_color = "#FFFFFF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  text_color = "#000000",
  text_font_size = "1.0rem",
  header_h1_font_size = "2.0rem",
  colors = c(myblue = "#006DAE", mywhite = "#FFFFFF")
)
```

@@

.center[
# 📚Thesis structure📚
]

@@

@@

# 🔍Regression diagnostics🔍

@@

- While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**. **Minor defects** in the model are **unlikely** to have a **significant impact** and may be best disregarded for practical purposes.

@@

.center[
# 📜Liteature of regression diagnostics📜
]

### Plotting the data is the recommended method by experts for diagnosing model fits.

@@

.center[
.caption[
Fig. 3: _Residual plot of a simple linear regression._
]
]

@@

# 🤔False visual discoveries🤔

@@

If we embed the residual plot in a matrix of **null plots**, we will find the visual discovery is

>>>>

---
title: "Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision"
subtitle: ""  
author: 
  - "Weihao (Patrick) Li"
date: '`r format(Sys.Date(), format="%B %d, %Y")`'
output:
  xaringan::moon_reader:
    seal: false
    css: [xaringan-themer.css, mine.css]
    nature:
      countdown: 55000
      beforeInit: "scripts/my_macro.js"
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
---

@@

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(tidyverse)
library(visage)
style_solarized_light(
  background_color = "#FFFFFF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i", "400i", "600"),
  code_font_google   = google_font("Fira Mono"),
  text_color = "#000000",
  text_font_size = "1.0rem",
  header_h1_font_size = "2.0rem",
  colors = c(myblue = "#006DAE", mywhite = "#FFFFFF")
)
```

@@

.center[
# 📚Thesis Structure📚
]

@@

#### Weihao Li, Dianne Cook, Emi Tanaka and Susan VanderPlas (2023)

.center[
# 🔍Regression Diagnostics🔍
]

**Residuals** summarise what is **not captured by the regression model**. Many different residual-based hypothesis tests are available to detect specific model defects.

<br>
<br>
<br>

.pull-right-33p[

Non-linearity: 
  - **F-test**
  - **RESET test** (Ramsey 1969)
]

.pull-right-33p[

.pull-right-33p[

@@

# 🔍Regression Diagnostics🔍

@@

- While such rejections may be **statistically correct**, their **sensitivity** may render the results **impractical**. **Minor defects** in the model are **unlikely** to have a **material impact** and may be best disregarded for practical purposes.

@@

.center[
# 📜Liteature of Regression Diagnostics📜
]

### Grahpical approaches (plots) are the recommended methods for diagnosing model fits.

@@

.center[
.caption[
Fig. 3: _Residual plot of a simple linear regression (more meaningful)._
]
]

@@

# 🤔False Visual Discoveries🤔

@@

# 🔬Visual Inference🔬

<br>
<br>
<br>

If we embed the residual plot in a matrix of **null plots**, we will find that it is **not uncommon** for residual plots to exhibit a triangle shape.

- The visual pattern is **over-interpreted**!

### The visual discovery is calibrated via comparison.

# 🔬Visual Inference🔬

**Visual inference** was introduced by Buja, et al. (2009) as an inferential framework to extend confirmatory statistics to visual discoveries.

A **lineup** consists of $m$ randomly placed plots, where one plot is the **data plot** and the remaining $m − 1$ plots (**null plots**) contain data consistent with the null hypothesis.

To perform a **visual test**, observer(s) will be asked to select the **most different plot(s)** from the lineup.
